---
title: "p8105_hw1_tl3195"
author: "Tingyi Li"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem 1

Load the moderndive library:

```{r}
library("moderndive")
```

load the early_january_weather dataset:
```{r}
data("early_january_weather")
```

Description of early_january_weather dataset:

The dataset contains EWR's weather conditions from Jan 1st to Jan 15th in 2013. The weather conditions were observed every hour and there are verious weather-related variables including temperature, dewpoint, humidity, wind direction, wind speed, wind gust, precipitation, pressure, and visbility.
The dataset has `r nrow(early_january_weather)` observations from January 1st to January 15th and `r ncol(early_january_weather)` weather_related variables. 
The mean temperature in this dataset is `r mean(early_january_weather$temp, na.rm = TRUE)` Fahrenheit.

Make a scatterplot of temp (y) vs time_hour (x); color points using the humid variable (adding color = ... inside of aes in your ggplot).

```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point()
```

Description of the plot:

The scatterplot indicates that there is an generally increasing trend in temperature when the time pass by. Moreover, the humidity also shows an increasing trend around January 13th.

Export your scatterplot to your project directory using ggsave.

```{r}
ggsave("p8105_prob1_scatterplot.pdf")
```


#Problem 2

Create a data frame (`prob2_df`) comprised of:

`sample`: a random sample of size 10 from a standard Normal distribution
`log_vec`: a logical vector indicating whether elements of the sample are greater than 0
`char_vec`: a character vector of length 10
`fac_vec`: a factor vector of length 10, with 3 different factor “levels”

```{r}
prob2_df <- tibble(
  sample = rnorm(10),
  log_vec = sample > 0,
  char_vec = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
  fac_vec = factor(c("small", "small", "small", "small", "medium", "medium", "medium", "large", "large", "large"))
)
```

Take the mean into the dataframe:

```{r}
mean(pull(prob2_df, sample))
mean(pull(prob2_df, log_vec))
mean(pull(prob2_df, char_vec))
mean(pull(prob2_df, fac_vec))
```

According to the results, a numeric and a logical variables work but a character and factor variables don't work.

```{r, results='hide', warning=FALSE}
as.numeric(pull(prob2_df, log_vec))
as.numeric(pull(prob2_df, char_vec))
as.numeric(pull(prob2_df, fac_vec))
```

For logical variables, TRUE is converted to 1 and FALSE is converted to 0. Hence, we can calculate the mean and the mean is the proportion of the TRUE values.
For character variables, when computing the mean, character variables are converted to NA. Hence, we cannot calculate the mean.
For factor variables, they have potential integer representation to represent different levels. When applying as.numeric, it's possible for them to be converted to 1, 2, and 3, respectively. However, the mean might not be meaningful as they cannot be interpreted.